### 2014-12-31 Studies/2013_PG_County/Report/PG_Co_Fire.tex
PG Co: update report

The value for the heat of combustion of wood used in this simulation was~16,400 kJ/kg, based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the amount of energy per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

@@

To calculate the total fire size, the appropriate burning area of the basement needed to be determined. Based on the floor plan of the structure and post-incident images, two distinct fires were considered: 1 source in the basement bathroom and 1 source in the basement kitchen.

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The fires started at the beginning of the simulation and increased to a value of 8.5~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. Since the actual ignition time is unknown, the fire was ramped quickly to reflect the conditions determined from reports and post-incident images. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations in the basement (1 source for the kitchen, 1 source for the bathroom).  At 207~s, using a simple decay model, the fire size decreases lineraly to 25~\% of its peak value in 60~s. This is to account for the impact of firefighters putting water on the basement fire (Table~\ref{tab:fire_info}). The following lines define the source fires in the FDS input file:

@@

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condtion, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompasses the structure with a total volume 14~m (46~ft) by 12~m (39ft) by 8~m (26~ft) that required a total of 1.3 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. The 20~cm grid resolution extends the computational domain 5~m to the left and right sides and 4~m behind the 10~cm mesh to provide sufficient computational domain for the wind profile to develop around the structure. The mesh is coarser to cut down on computational time and because high accuracy in the fluid mechanics of the air flow far from the structure is not important in the analysis of this simulation. The 20~cm resolution resulted in an additional 216,000 computational cells.

@@

At the estimated time of firefighter entry to the first floor (142~s into the simulation; 21:14:42 according to Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. Velocities through both the doorway at the top of the stairs and the front door were approximately 13~m/s (29~mph), faster than the incident wind  speed (9~m/s or 20~mph) because of the addition of the bouyancy of the fire plume. The model temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750-800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against the front door 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.

>>>>

The value for the heat of combustion of wood used in this simulation was 16,400~kJ/kg, based on data provided in the SFPE Handbook~\cite{SFPE:Tewarson}. The heat of combustion quantifies the amount of energy per unit mass of the fuel. The following text defines Eq.~\ref{eq:wood_comb} for an FDS input file with the desired fuel and reaction properties discussed above:

@@

To calculate the total fire size, the appropriate burning area of the basement needed to be determined. Based on the floor plan of the structure and post-incident images, two distinct fires were considered: 1 source in the basement bathroom and 1 source in the basement kitchen. The bathroom fire was estimated to be comprised of wood paneling along the four walls (17.3~m$^2$ or 186~ft$^2$) and wood subfloors and studs along the ceiling (2.9~m$^2$ or 31~ft$^2$). This results in 20.2~m$^2$ (217~ft$^2$) of wood burning surface area at 50~kW/m$^2$ or a 1~MW energy release rate.

The remainder of the combustable load in the basement was lumped into the localized kitchen fire.

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifications were made. The fires started at the beginning of the simulation and increased to a value of 8.5~MW in 10~s. The 10-s ramp up time was used to prevent abrupt changes in the velocity at the fire boundary condition in FDS. Since the actual ignition time is unknown, the fire was ramped quickly to reflect the conditions determined from reports and post-incident images. The second assumption is that the sources of the gas phase wood fuel have constant areas at fixed locations in the basement (1 source for the kitchen, 1 source for the bathroom).  At 207~s, using a simple decay model, the fire size decreases linearly to 25~\% of its peak value in 60~s. This is to account for the impact of firefighters putting water on the basement fire (Table~\ref{tab:fire_info}). The following lines define the source fires in the FDS input file:

@@

As discussed in Section~\ref{geom}, the structure has dimensions of 9.14~m (30~ft) by 7.62~m (25~ft) and a total height of 7.31~m (24~ft). To sufficiently resolve the wind boundary condition, the computational domain was extended beyond the volume of the structure. The large domain led to the use of two separate grid resolutions, a 10~cm (3.9~in) resolution and a 20~cm (7.8~in) resolution. The 10~cm resolution encompasses the structure with a total volume 14~m (46~ft) by 12~m (39ft) by 8~m (26~ft) that required a total of 1.3 million computational cells. As a result, all of the input model geometry lengths, ventilation openings (holes, doors, and windows), and fire areas will snap to the nearest 10~cm. The 20~cm grid resolution extends the computational domain 5~m to the left and right sides and 4~m behind the 10~cm mesh to provide sufficient computational domain for the wind profile to develop around the structure. The mesh is coarser to cut down on computational time and because high accuracy in the fluid mechanics of the air flow far from the structure is not important in the analysis of this simulation. The 20~cm resolution resulted in an additional 216,000 computational cells.

@@

At the estimated time of firefighter entry to the first floor (142~s into the simulation; 21:14:42 according to Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. Velocities through both the doorway at the top of the stairs and the front door were approximately 13~m/s (29~mph), faster than the incident wind  speed (9~m/s or 20~mph) because of the addition of the buoyancy of the fire plume. The model temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750-800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against the front door 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.
